\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{The Pythagoreans}
\author{Eric Yee}
\date{February 2024}

\begin{document}

\maketitle

I've dipped into a stream of divine teachings, bathing in the calm, fed the word of PATHAGORAS. For 5 years I have kept my vow of silence. Now my initiation is over. 

I have reached the plaza of Croton. The Pythagoreans - no - my brothers wait for me inside the temple. I ascend, an ethereal breeze fills my toga, and I feel as if I'm lifted to the gateway. My throat feels dry, I panic, do I even remember how to speak? But I feel Pythagoras's ideals empower me, my soul has been purified, rid of corruptions, and with enlightenment, I declare myself the newest member of the Pythagoreans!

Today is a day I long dreamt about, to follow in the ways of Pythagoras himself. The man who was the first foreigner to partake in Egyptian worship, who has undergone initiations into the Eleusinian Mysteries, learned the secrets of geometry from the magi of Persia.

But most of all I want to transcend. Pythagoras was said to have a golden leg, appearing simultaneously in Samos and Croton, he is as close to god as I can find in this mortal realm. Moreover, he has espoused the truth of reincarnation, and that he can remember all his past lives, starting with his first as the son of the Olympic god Hermes who had gifted him this supernatural ability. 

However, the divine notions were not enough to make me undergo the arduous initiation process. No, it was his compelling theories of numbers. The Pythagoreans presented a logical theory that drew me in. In their teachings, they explained to me that all ideas and concepts were just numbers. Number 1 meant intelligence, unity, everything! Number 2 was our ideas. And the number 3 was the most important. It was the Triad. The beginning, the middle, the end.

Then as if proving that he had discovered the divine truth, Pythagoras was able to fit music and harmonies into his idea of mathematical truth. Converting notes into quantifiable numerical intervals.

Ecstatic I have poured myself into the pious lifestyle of the Pythagoreans. It is a demanding and devout life-style. I had to give up my love of meat and be vegetarian. I must walk on grasses and fields since I am not allowed to use public roads. But one sacrifice was almost too much to bear. I could never eat BEANS! I would never again taste the savory salts of chili, hummus dip for my Doritos, and most of all my 100\% soy milk!

But alas, the price to transcend is worth the sacrifice. See you on the other side.

\end{document}
