\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amssymb}
\usepackage{amsmath}

\title{Polya's Neverending Walk}
\author{Eric Yee}
\date{November 2023}

\begin{document}

\maketitle

\section{Introduction}
It was a long time ago when I was safe at home. Back then, in the warm embrace of the origin, I was sheltered and safe, yet I longed for more--sigh--if only I had known better... 

Not that anyone will be able to read it. This a documentation of my journey.

It all started a long time ago, too long to remember. I had grown up in the loving embrace of my hometown: the Origin. It was all I knew, but unlike the others in my town, I longed for more. 

I always thought \emph{there must be something outside this 0-dimensional cage}. Unfortunately, I was right.

One day, I stumbled onto a mysterious trail on the outskirts of the Origin. A sign stood guard:\\
\emph{
The Number Line: \\
Forwards backward, with equal chance,\\
Adventure waits at each advance.\\
}
Without thinking, I leaped onto the trail. Finally, a chance to journey beyond the 0 dimensions! 

Bounding up and down, I raced across the Number Line, and before I knew it, I came upon a clearing. I looked around and I found myself in the center of another point, not unlike the origin. \emph{There were more points in this world?} In the distance I could make out the entrance I had run through, and in the opposite direction, an exit... or was that the entrance and this the exit... or... had I forgotten? 


No worries, I love mystery. I randomly chose an exit and kept leaping down the number line.  Once again, in a few seconds, I was at the center of another empty point. And yet again, there was an entrance and exit, but which was which? Again, I was disoriented. But, never discouraged, I chose randomly and journeyed on and on...

By now many of you might have ascertained the nature of my walk. I was on the 1-dimensional Number Line (you might know it as $\mathbb{Z}$), and, as I traversed from one integer to a consecutive one, I was forced to choose randomly to go forward or backward. Being one of the brightest back at the Origin, I figured this out as well, and my mathematical intuition also gave me a peculiar feeling that I must be destined to return to the Origin. 

Soon I proved this claim: Let $p$ be the probability I return to the origin if I start at the origin, ie. $p$ is the probability you reach the origin if you start at $x=1$.
\begin{align*}
    p &= P\left(\text{Return to the origin from $x=1$}\right)\\ 
    &= \frac{1}{2} + \frac{1}{2}P\left(\text{Return to the origin from x=2}\right)\\
    &=\frac{1}{2} + \frac{1}{2}\left(\text{Go to $x=1$ from $x=2$}\right) \cdot P\left(\text{Return to the origin from $x=1$}\right)\\
    &=\frac{1}{2} + \frac{1}{2}p\cdot p\\
    \Rightarrow     (p-1)^2&=0\\
    \Rightarrow          p&=1
\end{align*}
In other words: After I have taken my first step to $x=1$, I can either go backward and return to the origin with a $\frac{1}{2}$ probability or take a step forward to $x=2$ with a $\frac{1}{2}$ probability. To then return to the origin from $x=2$ means you have to arrive at $x=1$ at some point. Starting at $x=2$ and returning to $x=1$ is akin to starting and returning to the origin. From $x=1$, returning to the origin has a probability of $p$. Thus we solve our quadratic and find $p=1$


Indeed, my calculations did not fail me, after passing $8$ points, I found myself back at the Origin. But now that I knew there was a world outside the origin, my desire to see more of the world was insatiable. I would make journey after journey on the number line, and every time I would find myself back at the Origin without fault.

Eventually, I longed for more, and that was when I stumbled on \emph{the Plane}. 

I had just returned from a walk on the number line, and while strolling around the Origin I found myself in a circular clearing. To my front I saw another numberline-like trail stretching in front of me. Turning 90 degrees I saw an identical one, and another one in 180 degrees, and another in 270 degrees. Underneath my feet was a carving in the stone that read:

\emph{The Plane\\
Each direction, not two but four\\
Four ways to wander, no need to ponder
}\\

Immediately I understood where I was. While such an object could never exist in my $0$-dimensional home, I had long conjectured the existence of a $2$-dimensions that was like a combination of two \emph{Numberline}'s. Once again I was ecstatic. Running down a random direction I began my exploration of the \emph{Plane}, and much like last time at every new point there were $4$ directions I must choose randomly.

As I walked, a thought soon struck me. I had passed by thousands of points and I still had not returned to the origin. Maybe I might never return? My worries were unfounded, I soon worked out that I had a $100$\% chance of returning to the origin. Indeed after a few hundred more points, I was back in the origin. 

Now you might guess where I found myself next. Yes, a few days later I found myself in the center of a sphere, and in all $6$ directions - up, down, left, right, forward, and backward- I saw a numberline-like path extending outwards. 
\emph{The Hyperplane\\
With just one more dimension\\
Forever travel in random motion
}

I bounded upwards and found myself at another sphere/point with 6 exits. I ran down one, I don't know which, and repeated, and repeated, and repeated... I had again gone past a thousand points before the same existential dread crept up on me. \emph{Can I ever return?} I brushed it aside. I had a unity chance of returning on the \emph{Numberline} and the \emph{Plane}, why should it be different for the \emph{Hyperplane}? 

I kept going and with every point I passed, the feeling returned bit by bit. After passing another thousand points, I began to panic. \emph{What if I could never return}. 

By now I had already worked out the truth. Desperately, I had been trying to prove that I had a $1$ probability of returning. I failed every time, and soon I realized I could prove it false. 

I could prove it FALSE. I may never return... 

Memories of my beautiful hometown washed over me: The gentle hug of $0$-dimensions, the sublime beauty of nothing, the warm embrace of the Origin. In my mind's eye, the Origin had completely transformed. It wasn't a cage. I wasn't locked inside. Everything else was locked outside. And now I was locked outside...

I now knew that the probability of returning was infinitely slim, and with every passing point, it grew smaller. The poem that accompanied the \emph{Hyperplane} now made sense. I was destined to wander forever.

And so I have, and so I will be. 


\end{document}
